---
title: "Dynamic Documents in R"
author: "Andrew Duncan"
date: ""
output: 
  ioslides_presentation:
    css: sruc.css
    fig_height: 5
    fig_width: 7
    highlight: haddock
    widescreen: true
    logo: Images/SRUC.png
---

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = F, eval = T, comment = NA, warning = F, message = NA)
knitr::knit_hooks$set(wrapper = function(before, options, envir) {
  if (before) {
    sprintf('    ```{r %s}\n', options$params.src)
  } else '    ```\n'
})
```

## Our Normal Workflow

- Load data into R.
- Analyse data in R.
- Produce output results (table/plot)
- Type everything up in Word (maybe tables in Excel).
- There might be an easier way.


## Dynamic Documents in R

- Easily produce Word, PDF, HTML and presentations from inside R. 
- No copying and pasting images.
- Include code or not, it is up to you.
- Want to run the same analysis on different data - just change the input data. 


## Reproducible Research

- One of the central ideas of this is reproducible research. 
- Word and Excel normally involve a lot of pointing and clicking - hard to reproduce.
- Introducing R Markdown.
- Possible to write entire LaTeX documents from within R but LaTeX much harder to learn thatn R Markdown.


## R Markdown

- Developed by R Studio - works with base R too.
- Takes markdown (type of plain text document) and adds a lot of extras.
- **Knits** document into your chosen format. Using `knitr` package.

- We'll run through a few basics of markdown and then see some examples.

## Markdown Basics

- Markdown lets you do a lot of what you would do in Word. 
- Uses special syntax.
- For example: 
    + Bold `**this is bold text**` **this is bold text**
    + Italic `*this is italic text*` *this is italic text*
    + Superscript `R^2^` R^2^
    + Subscript `x~2~` x~2~
    + LaTeX `$$\int_1^2 \frac{1}{x} \: \mathrm{d}x$$` $$\int_1^2 \frac{1}{x} \: \mathrm{d}x$$


## R in Markdown

- Can also add R code directly into document (including plots) either stand alone or inline. 
- The same goes for plots. We can include the R code in the output if we wish.

```{r, echo = T, eval = F}
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point(aes(colour = factor(cyl))) + 
  geom_smooth() + 
  guides(colour = guide_legend(title = "Cylinder"))
```

## R in Markdown

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point(aes(colour = factor(cyl))) + 
  geom_smooth() + 
  guides(colour = guide_legend(title = "Cylinder"))
```

## Rmarkdown Examples

- I'll work through a couple of examples in RStudio first then base R.
- Note: 
    + To make a PDF you need LaTeX installed (Open Source). 
    + To make a Word doc you need Word/Libre Office installed.


## Compile Notebooks

- If you've just been plain old R code - you can make documents too.
- Include a YAML header and particular types of comments.
- Here's an example.


## Summary

- For Rmarkdown create an `.Rmd` file. 
- Edit to your heart's content.
- Click `knit` button in RStudio or
```{r ,eval = F, echo = T} 
rmarkdown::render(input = "my_file.Rmd", output_format = "word_document")
```
- There is a lot more that can be done ...
- Any questions?