---
title: "Dynamic Documents in R"
author: "Andrew Duncan"
date: ""
output: 
  ioslides_presentation:
    css: sruc.css
    fig_height: 5
    fig_width: 7
    highlight: haddock
    widescreen: true
    logo: Images/SRUC.png
---

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = F, eval = T, comment = NA, warning = F, message = NA)
knitr::knit_hooks$set(wrapper = function(before, options, envir) {
  if (before) {
    sprintf('    ```{r %s}\n', options$params.src)
  } else '    ```\n'
})
```

## Our Normal Workflow

- Load data into R.
- Analyse data in R.
- Produce output results (table/plot)
- Type everything up in Word (maybe tables in Excel).
- There might be an easier way.

## Reproducible Research

- Make your work reproducible (for yourself and for others). 
- Word and Excel normally involve a lot of pointing and clicking - hard to reproduce.
- LaTeX might be overkill. Can't run R in LaTeX editor.
- Maybe R Markdown would be of use.

## R Markdown

- Easily produce Word, PDF, HTML and presentations from inside R. 
- No copying and pasting images. 
- Include your code (or don't - it is up to you).
- Easily run the same analysis on multiple datasets (formatted the same).


## R Markdown

- Developed by R Studio - works with base R too.
```{r install_packages, echo = T, eval = F}
install.packages("rmarkdown")
library(rmarkdown)
```
- Takes markdown (type of plain text document) and adds a lot of extras.
- **Knits** document into your chosen format. Using `knitr` package.

- We'll run through a few basics of markdown and then see some examples.

## Markdown Basics

- Markdown lets you do a lot of what you would do in Word. 
- Rather than point and clicking uses special syntax.
- For example: 
    + Bold: `**this is bold text**` **this is bold text**
    + Italic: `*this is italic text*` *this is italic text*
    + Superscript: `R^2^` R^2^
    + Subscript: `x~2~` x~2~
    + LaTeX: `$$\int_1^2 \frac{1}{x} \: \mathrm{d}x$$` $$\int_1^2 \frac{1}{x} \: \mathrm{d}x$$


## R in Markdown

- Can also add R code directly into document (including plots) either stand alone or inline. 
- The same goes for plots. We can include the R code in the output if we wish.

```{r, echo = T, eval = F}
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point(aes(colour = factor(cyl))) + 
  geom_smooth() + 
  guides(colour = guide_legend(title = "Cylinder"))
```

## R in Markdown

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point(aes(colour = factor(cyl))) + 
  geom_smooth() + 
  guides(colour = guide_legend(title = "Cylinder"))
```

## Rmarkdown Examples

- I'll work through a couple of examples in RStudio first then base R.
- Note: 
    + To make a PDF you need LaTeX installed (Open Source). 
    + To make a Word doc you need Word/Libre Office installed.

## Summary

- For Rmarkdown create an `.Rmd` file. 
- Edit to your heart's content.
- Click `knit` button in RStudio or
```{r ,eval = F, echo = T} 
rmarkdown::render(input = "my_file.Rmd", output_format = "word_document")
```
- Just scratched the surface - <a href = "http://rmarkdown.rstudio.com/" target = "_blank">http://rmarkdown.rstudio.com/</a>.
- All notes and documents available on Github  - <a href = "https://github.com/aj2duncan/dyn_docs_pres"> https://github.com/aj2duncan/dyn_docs_pres </a>
- Any questions?