---
title: "Demonstration Dynamic Document"
author: "Andrew Duncan"
date: ''
output:
  html_document:
    code_folding: hide
    fig_height: 6
    fig_width: 7
    toc: yes
    toc_float: yes
  pdf_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # should the R code used be printed
                      eval = TRUE, # should the R code be evaluated
                      comment = NA, # Nothing preceding results
                      message = F, # Don't print any messages from R
                      warning = F) # Don't print any warnings from R

# Variables to change
my_fav_team = "NYY" # my favourite team - we'll change this to NYY later 
num_hr = 300 # Number of home runs to examine later
```

# Let's Play Ball

## Batting Average

To demo reproducible documents I decided to investigate some baseball statistics. 
First we load the data and any other packages that we want. We have supressed any warnings or comments. 

```{r loading_packages_data}
library(Lahman) # Contains baseball data
library(dplyr)
library(ggplot2)

# Calculate batting statistics
Batting = tbl_df(battingStats()) 
Teams = tbl_df(Teams)
Teams = select(Teams, yearID, teamID, franchID)
Batting = left_join(Batting, Teams)
team_name = TeamsFranchises$franchName[TeamsFranchises$franchID == my_fav_team]
```

Then we choose to examine my favourite team - the `r team_name`. 

```{r choose_team}
my_team = Batting %>%
  filter(franchID == my_fav_team)
```


Let us examine their batting over their entire history. 

```{r plot_ba}
ba_year = my_team %>%
  group_by(yearID) %>%
  summarise(average_ba = mean(BA, na.rm = T)) %>%
  ungroup()

ggplot(data = ba_year, mapping = aes(x = yearID, y = average_ba)) + 
  geom_point() + 
  xlab("Year") + ylab("Mean Batting Average") + 
  ggtitle(paste("Mean Batting Average of the", team_name, sep = " "))
```

## More than `r num_hr` Home Runs

We can tabulate the players who have hit `r num_hr` or more home runs for the `r team_name`. Note this does not include home runs hit for other teams - just the `r team_name`.

```{r collate_home_runs, results = 'asis'}
# Gather the players 
home_runs = my_team %>%
  group_by(playerID) %>%
  summarise(total_hr = sum(HR)) %>%
  filter(total_hr >= num_hr) 

# collect actual player names
players = tbl_df(Master) %>% 
  select(playerID, nameFirst, nameLast)

# join dataframes together and output
home_runs = left_join(home_runs, players) %>%
  mutate(Player = paste(nameFirst, nameLast, sep = " ")) %>%
  select(Player, total_hr) %>% # reordering columns
  arrange(desc(total_hr))

# print table
knitr::kable(home_runs, col.names = c("Player", "Total Number of Home Runs"))
```

